---
title: "S1. Supplemental Materials"
author: "Robinaugh et al."
output:
  pdf_document: default
  html_document: default
  word_document: default
---

S1. Supplemental materials for Robinaugh et al., Computer-Based Naming Treatment for Semantic Variant Primary Progressive Aphasia and a History of Traumatic Brain Injury: A Single-Case Experimental Design

```{r setup, include=F}
library(here)
library(knitr)
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

```{r packages, warning=F, message=F, include=F}
# load packages
library(kableExtra)
library(tidyverse)
library(brms)
library(cmdstanr)
library(tidybayes)
library(bayesplot)
library(ggthemes)
library(bayestestR)
library(janitor)
library(lme4)
```

```{r reading in data, include=F}
df <- read.csv(here::here("Data", "07_Language", "Robinaugh", 'SDTBI001_Data.csv')) %>%
  janitor::clean_names() %>%
  mutate(condition = ifelse(condition == 0, 'untreated', 'treated'),
         phase = as.factor(phase)) %>%
  mutate_if(is.character, as.factor) 

```

# Introduction
 
This document includes all posterior estimates for population-level and group-level effects for the Bayesian Mixed-Effects models described in the manuscript (Robinaugh et al., "Computer-Based Naming Treatment for Semantic Variant Primary Progressive Aphasia and a History of Traumatic Brain Injury: A Single-Case Experimental Design") and further detailed in Appendix A and Appendix B.

# Bayesian generalized mixed-effects model: naming performance

Naming performance over the course of treatment was evaluated using an Interrupted Time Series approach and Bayesian mixed effects modeling of item-level responses (correct or incorrect) with population-level effects (i.e. fixed effects) of intercept, baseline slope, level change, and slope change. Models included group-level intercepts for items nested within sets with corresponding group-level slopes for the primary three population-level effects, allowing each item and set to have their own intercept (an estimate of performance prior to the first baseline session), baseline slope, level change, and slope change.

## Interrupted Time Series Model

```{r model setup, message = F, cache = T, include = F}
df_coded2 <- df %>%
  filter(session < 13) %>%
  group_by(id) %>%
  mutate(phase = ifelse(phase == 'Pre', 0, 1),
         num_baselines = sum(phase == 0),
         # slope change is calculated by (T-(n1 + 1))D
         slope_change = (session - (num_baselines+1))*phase
         ) %>%
  select(id, set, condition, response, session, phase, slope_change) %>%
  mutate(phase = ifelse(condition == "untreated", 0, phase),
         slope_change = ifelse(condition == "untreated", 0, slope_change))

```

```{r model fitting,  message = F, cache = T, echo = T, results = 'hide'}
# Model Fitting
# Weakly informative priors assist model estimation

m.1b <- brm(response ~ 0 + Intercept + session + phase + slope_change + 
             (session + phase + slope_change | set/id), # 
           family = bernoulli(),  #binomial in lme4
           data = df_coded2, # both trained and untrained sets
           iter = 4000, # number of total samples per chain
           warmup = 1000, # number of samples tossed out per chain
           inits = 'random', # each chain takes random starting values
           chains = 4, # 4 total chains
           cores = 4, # use 4 computer cores to run chains simultaneously
           control = list(adapt_delta = .97), # to aid model convergence
           prior = c(prior(normal(-1, 2), class = b, coef = Intercept),
                     prior(normal(0, 2), class = b), # prior distributions
                     prior(normal(0, 2), class = sd),
                     prior(lkj(2), class = cor)),
           backend = 'cmdstan',
           seed = 42,
           refresh = 0, #for the markdown document 
           file = "new_approach",
           file_refit = "on_change"
)

```

## Interrupted Time Series Results

```{r summary, message =F, cache =T}

model_summary_with_posteriors <- function(model_name){
posterior_samples(model_name)[,1:4] %>%
  summarize(across(.fns = function(x){val = mean(ifelse(abs(x)>0.18,0,1))})) %>%
  t() %>%
  as_tibble() %>%
  rename(posterior_probability = V1) %>%
  bind_cols(as_tibble(fixef(model_name, probs = c(0.05, .95)),
                      rownames='Parameter')) %>%
  select(Parameter, Estimate, Est.Error, Q5, Q95, `ROPE %` = posterior_probability) 
} 

 model_summary_with_posteriors(m.1b) %>%
  kable(digits = 2, caption = "M.1b model summary")
```

## Effect Size Estimate of Trained Words

```{r effect size estimate, message = F, cache = T, echo = T}
pred_d <- m.1b$data %>%
  select(-response, -Intercept) %>% 
  group_by(id) %>% # for each word
  mutate(num_baselines = ifelse(set != 7, sum(phase == 0), 4)) %>% 
  #Untrained Set 7 set to give posterior distributions at session 4 
  #(last baseline without any tx) and session 12 (the last tx phase probe)
  filter(session == num_baselines | session == max(session))

es <- add_epred_draws(m.1b, newdata = pred_d |> filter(set != 7), 
                      pred = 'value', seed = 42) %>% 
  ungroup() %>% #
  mutate(time_point = ifelse(slope_change == 0, 'entry', 'exit')) %>% #
  select(time_point, id, value = .epred, draw = .draw) %>% #
  group_by(draw, time_point) %>% #
  summarize(num_corr = sum(value)) %>% #
  pivot_wider(names_from = time_point, values_from = num_corr) %>% # 
  mutate(effect_size = exit - entry) %>% #
  ungroup() %>%  # 
  select(effect_size)

es %>% # only need one variable
  tidybayes::median_qi(effect_size)

```

## Effect Size Estimate of Untrained Words

```{r effect size estimate untreated, message = F, cache = T, echo = T}
es_gen <- add_epred_draws(m.1b, newdata = pred_d |> filter(set == 7), 
                          pred = 'value', seed = 42) %>% 
  ungroup() %>% #
  mutate(time_point = ifelse(session == 4, 'entry', 'exit')) %>% # 
  select(time_point, id, value = .epred, draw = .draw) %>% #
  group_by(draw, time_point) %>% #
  summarize(num_corr = sum(value)) %>% #
  pivot_wider(names_from = time_point, values_from = num_corr) %>% # 
  mutate(effect_size = exit - entry) %>% #
  ungroup() %>%  # 
  select(effect_size)

es_gen %>% # only need one variable
  tidybayes::median_qi(effect_size)
```

# Bayesian generalized mixed-effects model: maintenance of trained words
## Trained words

Maintenance of gains (for trained words only) was evaluated using a generalized mixed-effects model with a four-level categorical population-level effect of timepoint (post-treatment, three month, six month, and 12 month follow-ups). Post-treatment was used as the reference value. Words were nested into sets and included as a group-level effect to reflect the structure of the data set, which includes six sets of trained words introduced to treatment at different time points during treatment. See Appendix B for a more detailed description of the model and analysis.

```{r maintenance setup, message = F, cache = T, eval = T, include =F}
df.mod3 <- df %>%
  filter(phase == 'Post' | phase == 'ThreeMo' | phase == 'SixMo' | phase == 'TwelveMo') %>% # only post and threemo timepoints
  droplevels() %>%
  mutate(phase = as.factor(phase), # phase as factor
         # condition as a factor with labels
         condition = as.factor(condition)) 

# post as reference 
contrasts(df.mod3$phase)
```

``` {r postthree, message = F, echo = T, cache = T, eval = T, echo = T, results = 'hide'}
m.3 <- brm(response ~ 0 + Intercept + phase +
             (1|set/id) +
             (session|id),
           family = bernoulli(),  #binomial in lme
           data = df.mod3 %>% filter(condition == 'treated'),
           iter = 5000, # number of total samples per chain
           warmup = 1000, # number of samples tossed out per chain
           inits = 'random', # each chain takes random starting values
           chains = 4, # 4 total chains
           cores = 4, # use 4 computer cores to run chains simultaneously,
           control = list(adapt_delta = .95), # to aid model convergence
           prior = c(
                     prior(normal(-1, 2), class = b, coef = Intercept),
                     prior(normal(0, 2), class = b), # prior distributions!
                     prior(normal(0, 2), class = sd),
                     prior(lkj(2), class = cor)),
           backend = 'cmdstan',
           seed = 42,
           refresh = 0 #this is only for the markdown document. 
)


```

```{r maintenance summary, message = F, echo = T, cache = T}
summary(m.3, prob = 0.90)
```

```{r session info}
sessioninfo::session_info() 
```

