---
title: "Open & Synthetic Data"
output:
  html_notebook:
    toc: yes
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

# Packages
Here, we load the packages that we will use for data wrangling, as well as the `synthpop` package, which we will use to generate synthetic data.
```{r}
library(tidyverse)
library(synthpop) # install.packages('synthpop')
```

# Data Synthesis {.tabset}
## 01 - Swallowing
This data was reported in "Normative Reference Values for FEES and VASES: Preliminary Data From 39 Nondysphagic, Community-Dwelling Adults" (https://pubs.asha.org/doi/full/10.1044/2023_JSLHR-23-00132) and was retrieved from https://osf.io/4anzm/.

### Loading Data
```{r}

```

### Generating Synthetic Data

### Comparisons

## 02 - Articulation
This data was reported in 

### Loading Data
```{r}

```

### Generating Synthetic Data

### Comparisons

## 03 - Fluency

## 04 - Voice

## 05 - Hearing

## 06 - AAC

## 07 - Language

## 08 - Cognition

## 09 - Social Communication

## {- .leave-tabset}