---
title: "Fluency"
author: Austin Thompson
---

This script recreates the analysis presented in Elsherif et al. (2021)

# Packages
```{r}
library(tidyverse)
library(effsize) # devtools::install_github("mtorchiano/effsize")
library(effectsize)
```

# Loading in the Data
Retrieved the data from: https://osf.io/q2b7v/ 
```{r}
data_fluency <- rio::import(file = "data_Dyslexia Stutter Master.csv") |> 
  filter(Rejection == 0) |> 
  filter(Group != "AWD")

data_fluency |> 
  group_by(Group) |> 
  summarise(
    N = length(Group),
    M = mean(Nonwordrepetition),
    SD = sd(Nonwordrepetition),
    SEM = sd(Nonwordrepetition) / sqrt(length(Nonwordrepetition)),
    CIneg = M - 1.96 * SEM,
    CIpos = M + 1.96 * SEM
  )
```

# Analysis
Here, we recreate the comparison between neurotypical and stuttering in non-word repetition, (t(36.42) = 6.61, p < .001, Δ = 1.26).
```{r}

t_test <- stats::t.test(Nonwordrepetition ~ Group,
                        data = data_fluency,
                        paired = FALSE,
                        var.equal = FALSE)

delta <- effectsize::glass_delta(Nonwordrepetition ~ Group,
                                 data = data_fluency,
                                 correction = FALSE,
                                 ci = 0.95)


# Pasting it together to get the APA write-up
apa <- paste0(
  "t(",
  
  # DF
  df <- t_test[["parameter"]][["df"]] |>
    round(digits = 2),
  
  ") = ",
  
  # t statistic
  t <- t_test[["statistic"]][["t"]] |>
    round(digits = 2),
  
  ", p < .001, Δ = ",
  
  delta$Glass_delta |>
    round(digits = 2)
  )

apa

```

The delta is the same from their Rmd, but different from what they report in the manuscript :(

# Analysis #2
Here, we recreate the comparison between neurotypical and stuttering in Test of Word Reading Efficiency (TOWRE) sight word reading, t(35.28) = 3.46, p = .0014, Δ = 0.66
```{r}

t_test <- stats::t.test(sightTOWRE ~ Group,
                        data = data_fluency,
                        paired = FALSE,
                        var.equal = FALSE)

delta <- effectsize::glass_delta(sightTOWRE ~ Group,
                                 data = data_fluency,
                                 correction = TRUE,
                                 ci = 0.95)

# Pasting it together to get the APA write-up
apa <- paste0(
  "t(",
  
  # DF
  df <- t_test[["parameter"]][["df"]] |>
    round(digits = 2),
  
  ") = ",
  
  # t statistic
  t <- t_test[["statistic"]][["t"]] |>
    round(digits = 2),
  
  ", p < .001, Δ = ",
  
  delta$Glass_delta |>
    round(digits = 2)
  )

apa
```

Again, the delta glass is similar to their R file, but different that what they reported.
