---
title: "R Notebook"
output: html_notebook
---

This script recreates the analysis in King et al. (2022)

# Packages
```{r}
library(tidyverse)
library(janitor)
library(gtsummary)
library(labelled)
```

# Loading in the Data
Data downloaded from: https://osf.io/sx3nf/
```{r}
qualtrics_csv = function(file){
  data = read_csv(file, col_names = FALSE, skip = 3)
  names = read_lines(file, n_max = 1) |> 
    str_split(pattern = ",") |> 
    unlist()
  names(data) = names
  labels = read_csv(file, col_names = FALSE, skip = 1, n_max = 1)
  labels = t(labels)
  labels = labels[,1]
  labels = setNames(as.list(labels), names)
  data = set_variable_labels(data, .labels = labels, .strict = FALSE)
  data = clean_names(data)
  data
}

df = qualtrics_csv(file = "data_Combined CSV data.csv") |> 
  janitor::remove_empty("cols") |> 
  janitor::remove_empty("rows") |> 
  filter(progress > 98) |> 
  select(start_date, participant_id, q2:x2f)
```

# Analysis
Here, we recreate the finding that for intervention both lack of/limited Internet and technology barriers increased significantly, χ2(2) = 34.6, p < .001
```{r}

chi_squared <- df |> 
  select(q87_1_1, q87_1_2, q87_1_3, q87_1_5, q87_1_6, q87_1_7,
         q87_2_1, q87_2_2, q87_2_3, q87_2_5, q87_2_6, q87_2_7,
         q87_3_1, q87_3_2, q87_3_3, q87_3_5, q87_3_6, q87_3_7,
         q89_1_1, q89_1_2, q89_1_3, q89_1_5, q89_1_6, q89_1_7,
         q89_2_1, q89_2_2, q89_2_3, q89_2_5, q89_2_6, q89_2_7,
         q89_3_1, q89_3_2, q89_3_3, q89_3_5, q89_3_6, q89_3_7) |> 
  pivot_longer(everything()) |> 
  separate(name, into = c("var", "time", "question"), sep = "_") |> 
  mutate(time = factor(time, labels = c("Prior to\nMarch 2020", "March to\nJuly 2020", "August 2020 to\nSpring 2021")),
         var = factor(var, labels = c("Assessment", "Intervention"))) |> 
  drop_na(value) |> 
  group_by(var, value) |> 
  summarize(
    prop = list(chisq.test(table(time)))
  ) |> 
  mutate(chi = map_dbl(prop, ~.x$statistic),
         df = map_dbl(prop, ~.x$parameter),
         p = map_dbl(prop, ~.x$p.value)) |>
  dplyr::filter(var == "Intervention") |>
  dplyr::filter(value == "Lack of/limited internet")

# Pasting it together to get the APA write-up
apa <- paste0(
  "χ^2(",

  # DF
  chi_squared$df,
  
  ") = ",
  
  # Chi Sqaured statistic
  chi_squared$chi |>
    round(digits = 2),
  
  ", p < .001"
  )

apa

```

We successfully recreated the finding! Woo!